---
title: "stl lead inequity"
format: html
editor: visual
---

## Day 1 Activities

Attach the tidyverse and janitor packages in a new code chunk

```{r}
library(tidyverse)
library(janitor)
library(skimr)
library(dplyr)
```

Read in the stl_blood_lead.csv data at stl_lead, use clean_name function to rename columns

```{r}
stl_lead <- read_csv(here::here("data",
                                "stl_blood_lead.csv"),
                     na = c(".")) %>% 
  clean_names()
```

Do some basic exploration of the dataset

```{r}
skim(stl_lead)
```

In a new code chunk, from `stl_lead` create a new data frame called `stl_lead_prop` that has one additional column called `prop_white` that returns the percent of each census tract identifying as white (variable `white` in the dataset divided by variable `totalPop`, times 100). You may need to do some Googling. Hint: `dplyr::mutate(new_col = col_a / col_b)` will create a new column `new_col` that contains the value of `col_a / col_b`

```{r}
stl_lead_prop <- stl_lead %>% 
  mutate("prop_white" = (white / total_pop) * 100)
```

Create a scatterplot

```{r}
stl_lead_plot <- ggplot(data = stl_lead_prop, aes(x = pct_elevated, y = prop_white))+
  geom_point(size = 3,
             alpha = 0.5)+
  theme_minimal()+
  labs(x = "Lead Blood Level", y = "White Popluation")

stl_lead_plot
```

Save to scatterplot to figs

```{r}
ggsave(plot = stl_lead_plot, here::here("figs", "stl_lead_test.png"),
       width = 6,
       height = 5)
```

Create a histogram

```{r}
pct_elevated_plot <- ggplot(data = stl_lead_prop, aes (x = pct_elevated))+
  geom_histogram(color = "cadetblue4",
                 fill = "cadetblue1",
                 bins = 15)+
  labs (x = "Lead Blood Level", y = "Count")

pct_elevated_plot
```

Save historgram to fig

```{r}
ggsave(plot = pct_elevated_plot, here::here("figs","pct_elevated_plot.png"),
       width = 6,
       height = 5)
```
